---
title: "Base R Learning"
author: "Meghan Fletcher"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# object and classes

## Vectors

### Creating a vector

```{r}
# Numeric vector
my_vector <- vector(mode = "numeric", length = 100)

# Logical vector
my_second_vector <- vector(mode = "logical", length = 20)
```

### Vector properties

#### Length

```{r}
# Length of my vector
length(my_vector)
length(my_second_vector)
```

#### Class

```{r}
class(my_vector)
class(my_second_vector)
```

#### Indexing (accessing) vectors

```{r}
# Using c to createa vector
my_vector <- c(1:50)

# What's the first element?
my_vector[1]

# What's the last element?
my_vector[-1]

# What's the seventh element?
my_vector[7]

# Select a range from a vector
my_vector[3:7]
```

## Matrices - vectors that have lengths in two directions

### Creating matrices

```{r}
# Create a 10 x 10 matrix full of NA's
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

### Matris properties

#### Dimensions

```{r}
# What are the dimensions? (gives number of rows then number of columns)
dim(my_matrix)
```

#### Length

```{r}
# How many elements are there?
length(my_matrix)
```

#### Class

```{r}
class(my_matrix)
```

#### Indexing (picking sometihng out)

```{r}
my_matrix[2,]

#Index a column
my_matrix[,3]

# Index a given cell
my_matrix[8,3]

# Index and replace
my_matrix[8,3] <- 100000

# Replace and entire column
my_matrix[,5] <- 10

```

# Control flow

## `for` loops

variable (counter), sequence, expression

```{r}
for (i in 1:10){
  print('hello world')
}
```

# Generating random nubers

```{r}
rnorm(n = 5, mean = 10, sd = 2)
```


# Let's combine all of the above

We will use `numeric` to create an empty vector
We will use `for` to iterate along items of that empty vector
We will use `[]` to access each slot and fill-in that empty vector
We will use `rnorm` to sample values from a random normal distribution
We will plot the trend

$$N_{t}=N_{t-1} + r\left(1 - \frac{N_{t-1}}{K}\right)$$

$$
r \sim norm(\mu = 1.5, \sigma = 0.8)
$$

## Writing the code down

```{r}
# Set up mode parameters
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# Set up model objects
N <- vector(mode = "numeric", length = nsteps)
N[1] <- 25

# Model it
for(t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <- N[t-1] + r * (1 - (N[t-1] / K))
}

# Visualize it
plot(N)

```



# Something more
We will use `matrix` to create an empty matrix (rows are time, columns are runs)
We will repeat the process

```{r}
# Set up mode parameters
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# Set up model objects -Modify
N <- matrix(data = NA, nrow = nsteps, ncol = 100)
N[1,] <- 25

# Model it
for(t in 2:nsteps){
  r <- rnorm(n = 100, mean = mu, sd = sigma)
  N[t,] <- N[t-1,] + r * (1 - (N[t-1,] / K))
}

# Visualize it
matplot(N, type = "l")
```

# Bring in your friends!

Conver it to things we know about! `ggplot2`

```{r}
# Load tidyverse
library(tidyverse)

# Ctreate data.frame
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

# Visualize it
ggplot(data = my_data, mapping = aes(x = time, y = N)) +
  geom_line(aes(group = run), size = 0.1, alpha = 0.2) +
  stat_summary(geom = "line", fun = mean, color = "red", size = 1) +
  theme_bw() +
  labs(x = "Time (years)", y = "Population size (N)")
```





































